% !TEX root = submission.tex
\documentclass[lettersize,journal]{IEEEtran}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}

\usepackage{amsmath,amsfonts}
\usepackage{amssymb} % math symbols
\usepackage{amsthm}  % theorem/proof environments (IEEEtran-compatible)
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{array}
\usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
\usepackage{textcomp}
\usepackage{stfloats}
\usepackage{url}
\usepackage{verbatim}
\usepackage{graphicx}
\hyphenation{op-tical net-works semi-conduc-tor IEEE-Xplore}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\usepackage{balance}

% --- Theorem-like environments ---
% Keep counters independent (Theorem 1, Proposition 1, Lemma 1, Assumption 1, etc.)
\newtheorem{theorem}{Theorem}
\newtheorem{proposition}{Proposition}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}
\newtheorem{condition}{Condition}
\newtheorem{assumption}{Assumption}
\newtheorem{definition}{Definition}
\theoremstyle{remark}
\newtheorem{remark}{Remark}

\begin{document}
\title{Early-Terminable Energy-Safe Iterative Coupling for Parallel Simulation of Port-Hamiltonian Systems}
\author{Qi Wei, Jianfeng Tao, Hongyu Nie, Wangtao Tan
\thanks{Manuscript received January 16, 2026; accepted ~~ ~~~~~~ ~~~~;Date of publication ~~ ~~~~~~ ~~~~; This work was supported by the National
Natural Science Foundation of China (Grant No. 52075320). 

The authors are with the School of Mechanical Engineering, Shanghai Jiao Tong University, Shanghai 200240, China. (e-mail: jftao@sjtu.edu.cn)

Digital Object Identifier
}}
 
\markboth{IEEE ROBOTICS AND AUTOMATION LETTERS,~Vol.~, No.~, ~2026}%
{Early-Terminable Energy-Safe Iterative Coupling for Parallel Simulation of Port-Hamiltonian Systems}

\maketitle

\begin{abstract}
Parallel simulation of robotic systems requires partitioning the dynamics into coupled subsystems. Finite-iteration coupling across the partition boundary can inject spurious energy, even when each subsystem is passive. We propose an early-terminable, energy-safe coupling interface for port-Hamiltonian subsystems based on Douglas--Rachford splitting in wave (scattering) coordinates. The wave-domain formulation reduces passivity to norm inequalities and coupling to orthogonality. Within this setting, the deep correspondence between monotone operator theory and discrete passivity can be exploited to construct a Douglas--Rachford inner iteration whose Fejér monotonicity provides algorithmic dissipation. 
Under passivity of the subsystem integrators and an impedance-tuning condition, the proposed method guarantees discrete passivity of the augmented storage for any finite inner-iteration budget and converges to the monolithic discretization as the budget increases. Experiments on a linear--Duffing coupled-oscillator benchmark confirm the guarantees at numerical roundoff ($10^{-14}$ in double precision), with RMS state error decreasing monotonically with the inner budget.

\end{abstract}

\begin{IEEEkeywords}
Simulation and Animation, Dynamics, Formal Methods in Robotics and Automation.
\end{IEEEkeywords}


\input{01_introduction}
\input{02_problem_formulation}
\input{03_iterative_coupling}
\input{04_passivity_and_convergence}
\input{05_experiment}
\input{06_discussion}




\bibliographystyle{IEEEtran}
\bibliography{IEEEabrv,references} % Include the new bib file


\vspace{-1.5\baselineskip}



\end{document}
